Seldin Report: Ssec Scores

### 2022-07-12 SeldinReport_fix.Rmd
Getting there

1. Report on Ssec scores from "Seldin Methods"
  a. Find distribution of Ssec values
  b. Find distribution of correlation p-values
  c. Find a null distribution of Ssec values

THIS WAS DONE WITH TISSUE EXCLUSIVE GENES

@@

Previously compared Secreted Genes to Tissue-Specific Secreted Genes. For both of these categories, they were compared to the full set of target tissue genes. There was not a very significant difference between the two peaks.  
Below are the graphs for Tissue-Specific Secreted Genes and Tissue-Specific Non-Secreted Genes. These were both compared to full set of target genes. These also do not show a significant difference between the Ssec scores of Tissue-Specific Secreted and Tissue-Specific Non-Secreted.

# Calculate Ssec values: tissue-specific -> full
#### TS ####
## BRAIN ##

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("TSNS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

with(compare, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
abline(a = 0, b = 1)

# Liver
h1 = cbind(rep("TSNS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

with(compare, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
abline(a = 0, b = 1)

The graphs below modify the comparison slightly. Instead of comparing the tissue specific secreted and tissue specific nonsecreted to the full dataset as target genes, our target genes are ALL tissue specific genes in the target tissue (secreted and non-secreted). This hopefully eliminates the overlap with the non-tissue-specific genes and doesn't limit the association to just secreted genes. There is a very slight difference in the density graphs for these. The Tissue-Specific Secreted have a VERY slightly higher Ssec score on average than the Tissue-Specific Non-Secreted.

## Comparing Tissue-Specific Secreted to Tissue-Specific Non-Secreted with reference group Tissue-Specific Target Genes

@@

#write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
#write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
#write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
#write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")

@@

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("TSNS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

@@

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

#with(compare.b, qqplot(val[trt == "TSNS brain"], val[trt == "TSS brain"]))
#abline(a = 0, b = 1)

# Liver
h1 = cbind(rep("TSNS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

@@

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

#with(compare.l, qqplot(val[trt == "TSNS liver"], val[trt == "TSS liver"]))
#abline(a = 0, b = 1)

@@

@@

```{r, echo = FALSE, message = FALSE}

#brain

wil.brain = coin::wilcox_test(val ~ trt, compare.b)

# liver

wil.liver = coin::wilcox_test(val ~ trt, compare.l)

```{r, eval = FALSE, echo = FALSE}
# checking secreted

checking = sec_annot[sec_annot$ensembl_gene_id == top_brain[4],] # extracellular matrix

# ENSRNOG00000005770: "The WNT signalling modulator, Wise, is expressed in an interaction-dependent manner during hair-follicle cycling."

```
A Wilcoxon-Mann-Whitney test shows no difference between the two peaks on either graph with p-values `r coin::pvalue(wil.brain) ` for brain and `r coin::pvalue(wil.liver)` for liver. 
We can try comparing the Tissue-Specific Secreted to the Ssec values of all of the origin tissue proteins. I need to create that dataset in the server; attempts to complete in chunks have been unsuccessful.

```{r, echo = FALSE}
## Pvalue for top and bottom Ssec scores, liver and brain
# brain
plotting = melt(brain.liv$p, id.vars = "refseq")
cheese =  as.character(plotting$Var1) %in% c(top_brain, bot_brain)
limit = plotting[cheese, ]

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="P-Value Density Curve: Top6 and Bottom6 Brain Genes", x = "P-Value", y = "Density")

# liver
plotting = melt(liv.brain$p, id.vars = "refseq")
limit = plotting[as.character(plotting$Var1) %in% c(top_liver, bot_liver),]

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="P-Value Density Curve: Top6 and Bottom6 Liver Genes", x = "P-Value", y = "Density")

The above two plots are the density distribution of p-values within the proteins with the highest and lowest Ssec scores. We can see that the top Ssec score proteins have a peak on the lower end of p-values, which makes sense given they have the highest Ssec scores. These graphs do not appear as well-defined as the graphs from previous reports where the all target tissue genes were used as the reference. 
[Attempted to put pvalues on log scale but lost the distinction between top and bottom Ssec scores; Tried to include a line for a null distribution, but was having trouble with ggplot]

```{r, echo = FALSE}
# Correlation coefficients top and bottom Ssec scores, liver and brain
# brain
magn = brain.liv$bicor

plotting = melt(magn, id.vars = "refseq")
limit = rbind(plotting[as.character(plotting$Var1) %in% top_brain,], plotting[as.character(plotting$Var1) %in% bot_brain,])

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="Magnitude Density Curve: Top6 and Bottom6 Brain Genes", x = "Correlation", y = "Density")

# liver
magn = liv.brain$bicor
plotting = melt(magn, id.vars = "refseq")
limit = rbind(plotting[as.character(plotting$Var1) %in% top_liver,], plotting[as.character(plotting$Var1) %in% bot_liver,])

ggplot(limit, aes(value, color = Var1)) + 
  geom_density() +
  labs(title="Magnitude Density Curve: Top6 and Bottom6 Liver Genes", x = "Correlation", y = "Density")

Using the same top and bottom proteins, the above graphs show the correlation from the bicor calculation. The curves that are wider at the bottom are the top proteins, indicating that the more significant proteins also have greater correlation coefficients.

Below, we have used methods outlined in Seldin (2018) to complete a permutation test for the Ssec values of particular proteins. We can see with the significant protein below, the observed value is far above the 95% confidence interval of the null distribution. Below that, a protein that had the lowest Ssec value is far BELOW the 95% confidence interval (0.974 to 0.996). What this means has yet to be determined.

```{r, echo = FALSE, eval = FALSE}
# annotation

anno_brain = sec_annot[sec_annot$ensembl_gene_id %in% top_brain,]
anno_liver = sec_annot[sec_annot$ensembl_gene_id %in% top_liver,]

```
```{r, eval = FALSE, echo = FALSE, warning = FALSE}

#permute strain names in target tissue array
# bicor> pval> Ssec val
# repeat 1000- compare observed value to mean of permuted values
target_protein = protein_liver
perm_test = matrix(NA, ncol = 1000, nrow = 174)
perm_test = as.data.frame(perm_test)
rownames(perm_test) = ts_Ssec_brain[,1]

@@

brain.liv = bicorAndPvalue(ts_protein_brain, target_protein, use = 'pairwise.complete.obs')

@@

Ssec_brain_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(protein_liver))) %>%
  dplyr::select(-score)
  #arrange(desc(Ssec))

perm_test[,i] = Ssec_brain_perm[,2]
} 
```

@@

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
perm_test = read.csv("Permutation_Seldin_Report.csv")
perm_test= fix.read.csv(perm_test)

test_top_b = perm_test[rownames(perm_test) == top_brain[1],]
true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == top_brain[1],]
hist(as.numeric(test_top_b), main = "Null Ssec Distribution for Top Gene: ENSRNOG00000021020", xlab = "Ssec Score")
abline(v = true, col = "red")

test_bot_b = perm_test[rownames(perm_test) == bot_brain[1],]
true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == bot_brain[1],]
hist(as.numeric(test_bot_b), main = "Null Ssec Distribution for Bottom Gene: ENSRNOG00000026419", xlab = "Ssec Score")
abline(v = true, col = "red")

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
write.csv(perm_test, "Permutation_Seldin_Report.csv")

@@

setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")

full_brain = read.csv("Full_Brain_asso_TS.TSNS_Ssec.csv")
full_liver = read.csv("Full_Liver_asso_TS.TSNS_Ssec.csv")

# Visualizations
#hist(Ssec_brain$Ssec)
#hist(ts_Ssec_brain$Ssec)

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("Full brain", times = length(full_brain$Ssec)), full_brain$Ssec)
h2 = cbind(rep("TSS brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("Full liver", times = length(full_liver$Ssec)), full_liver$Ssec)
h2 = cbind(rep("TSS liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Full v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

# compare secreted to non-secreted

#### Secreted ####
## BRAIN ##
target_liver = cbind(ts_protein_liver, tsns_protein_liver)

brain.liv = bicorAndPvalue(secreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

sec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_brain = Ssec_brain[Ssec_brain[,1] %in% colnames(ts_protein_brain),]
top_brain = head(sec_Ssec_brain[,1])
bot_brain = tail(sec_Ssec_brain[,1])

target_brain = cbind(ts_protein_brain, tsns_protein_brain)

liv.brain = bicorAndPvalue(secreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

sec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

#ts_Ssec_liver = Ssec_liver[Ssec_liver[,1] %in% colnames(ts_protein_liver),]
top_liver = head(sec_Ssec_liver[,1])
bot_liver = tail(sec_Ssec_liver[,1])

#### Not Secreted ####

## BRAIN ##

brain.liv.c = bicorAndPvalue(notsecreted_protein_brain, target_liver, use = 'pairwise.complete.obs')
# brain.liv$p

nsec_Ssec_brain = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_liver))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

liv.brain.c = bicorAndPvalue(notsecreted_protein_liver, target_brain, use = 'pairwise.complete.obs')
# brain.liv$p

nsec_Ssec_liver = data.frame(Gene_symbol = names(scores), score = scores) %>%
  #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
  mutate(Ssec = score / length(colnames(target_brain))) %>%
  dplyr::select(-score) %>%
  arrange(desc(Ssec))

####### Visualizations

#Brain
h1 = cbind(rep("Not Secreted brain", times = length(nsec_Ssec_brain$Ssec)), nsec_Ssec_brain$Ssec)
h2 = cbind(rep("Secreted brain", times = length(sec_Ssec_brain$Ssec)), sec_Ssec_brain$Ssec)

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

with(compare.b, qqplot(val[trt == "Not Secreted brain"], val[trt == "Secreted brain"]))
abline(a = 0, b = 1)

# Liver
h1 = cbind(rep("Not Secreted liver", times = length(nsec_Ssec_liver$Ssec)), nsec_Ssec_liver$Ssec)
h2 = cbind(rep("Secreted liver", times = length(sec_Ssec_liver$Ssec)), sec_Ssec_liver$Ssec)

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Specific Non-Secreted v. Tissue-Specific Secreted", x = "Ssec", y = "Density")

with(compare.l, qqplot(val[trt == "Not Secreted liver"], val[trt == "Secreted liver"]))
abline(a = 0, b = 1)

#brain

wil.brain = coin::wilcox_test(val ~ trt, compare.b)
wil.brain

# liver

wil.liver = coin::wilcox_test(val ~ trt, compare.l)
wil.liver

# :'(

# random brain secreted genes

sample(secreted_protein_brain, size = 4, replace = FALSE)

#ENSRNOG00000012655: extracellular matrix
# ENSRNOG00000053272: endo reticulum, extracellular space, cytoplasm, perinuclear region

checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000012655",]
checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000001806",]
checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000008519",]

checking = sec_annot[sec_annot$ensembl_gene_id == "ENSRNOG00000021020",]

# random liber secreted genes

sample(secreted_protein_liver, size = 4, replace = FALSE)

# ENSRNOG00000058444: nucleus, chromosome, secreted
# ENSRNOG00000048642: same

# QQ plot

@@

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"]))
abline(a = 0, b = 1)

>>>>

1. Calculate Ssec values using biweight midcorrelation for TES Protein-Coding Genes
  a. Plot the distribution of Ssec values, comparing TES to TENS
  b. Plot the QQ plots of Ssec values, comparing TES to TENS
  c. Find null distribution of Ssec values
2. Select TES protein-coding genes to evaluate via enrichment analysis in MsigDB.Rmd
  a. Plot target genes for the top TES protein-coding genes

This report will document the results of the fix implemented to the biweight midcorrelation calculation of Ssec scores. Liver and brain datasets were verified to be in the same strain order in this report.

## Density Plots

Below are the density plots comparing the distributions of tissue-exclusive, secreted (TES) protein-coding genes and tissue-exclusive, non-secreted (TENS) protein-coding genes for both brain and liver. 
The plots for brain are fairly similar, however there is a slight bump at the right hand side of the plot where TES protein-coding genes have a higher density.

@@

@@

write.csv(ts_Ssec_brain, "Target_TESec_Ssec_brain.csv")
write.csv(ts_Ssec_liver, "Target_TESec_Ssec_liver.csv")
write.csv(tsns_Ssec_brain, "Target_TENSec_Ssec_brain.csv")
write.csv(tsns_Ssec_liver, "Target_TENSec_Ssec_liver.csv")

@@

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("TENS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
h2 = cbind(rep("TES brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

@@

ggplot(compare.b, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")

with(compare.b, qqplot(val[trt == "TENS brain"], val[trt == "TES brain"]))
abline(a = 0, b = 1)

## Histograms

c1 = rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
c2 = rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")

sec = hist(ts_Ssec_brain$Ssec, plot = FALSE) 
nsec = hist(tsns_Ssec_brain$Ssec, plot = FALSE)

plot(nsec, col = c1) 
plot(sec, col = c2, add = TRUE)

sec = hist(ts_Ssec_liver$Ssec, plot = FALSE) 
nsec = hist(tsns_Ssec_liver$Ssec, plot = FALSE)

plot(nsec, col = c1) 
plot(sec, col = c2, add = TRUE)

The plots for liver differ more dramatically, likely due to the small sample size for TES protein-coding genes in the liver. We continue to see a small bump in the lower right that may be promising.

# Liver
h1 = cbind(rep("TENS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
h2 = cbind(rep("TES liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

@@

ggplot(compare.l, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")

with(compare.l, qqplot(val[trt == "TENS liver"], val[trt == "TES liver"]))
abline(a = 0, b = 1)

@@

qqplotAnnot = function(x, y,
                       probs=c(0.001, 0.01, 0.1, 0.9, 0.99, 0.999),
                       ...)
{
  qqplot(
    x,
    y,
    pch=16, cex=0.5,
    ...
  )
  abline(0, 1, col="red")

  q_y = quantile(y, probs=probs, na.rm=TRUE)
  q_x = quantile(x, probs=probs, na.rm=TRUE)

  points(q_x, q_y, col="grey", cex=0.6)
  text(q_x, q_y, label=paste0(probs*100, "%"), pos=1, cex=0.8, col="grey")
}

qqplotAnnot(tsns_Ssec_brain$Ssec, ts_Ssec_brain$Ssec,
            main = "QQ-plot, Liver x adipose",
            xlab = "Non-secreted factor Ssec (-ln p)",
            ylab = "Secreted factor Ssec (-ln p)"
)

qqplotAnnot(tsns_Ssec_liver$Ssec, ts_Ssec_liver$Ssec,
            main = "QQ-plot, Liver x adipose",
            xlab = "Non-secreted factor Ssec (-ln p)",
            ylab = "Secreted factor Ssec (-ln p)"
)

## QQ Plots

The following are the QQ plots comparing the quantiles of Ssec scores for TES protein-coding genes and TENS protein-coding genes. 
In the plot for brain, we can see that the points follow the line fairly closely, with three points slightly above the null line. In the plot for liver, there is a pretty substantial dip of points below the null line, with no points above the line.

```{r, echo = FALSE}
################### QQ PLOTS #########################

@@

## Permutation Test

Below, we have used methods outlined in Seldin (2018) to complete a permutation test for the Ssec values of particular proteins. We can see with the significant protein below, the observed value is far above the 95% confidence interval of the null distribution. Below that, a protein that had the lowest Ssec value is far BELOW the 95% confidence interval (0.974 to 0.996).

#permute strain names in target tissue array
# bicor> pval> Ssec val
# repeat 1000- compare observed value to mean of permuted values

permy = function(target_protein, tes_protein_ssec, tes_protein_protein)
{
  perm_test = matrix(NA, ncol = 1000, nrow = nrow(tes_protein_ssec))
  perm_test = as.data.frame(perm_test)
  rownames(perm_test) = tes_protein_ssec[,1]

@@

    brain.liv = bicorAndPvalue(tes_protein_protein, target_protein, use = 'pairwise.complete.obs')

@@

    Ssec_perm = data.frame(Gene_symbol = names(scores), score = scores) %>%
     #filter(Gene_symbol %in% sec_annot$ensembl_gene_id) %>%
     mutate(Ssec = score / length(colnames(target_protein))) %>%
     dplyr::select(-score)
     #arrange(desc(Ssec))

    perm_test[,i] = Ssec_perm[,2]
  }

  return(perm_test)
}

perm_test.b = permy(target_liver, ts_Ssec_brain, ts_protein_brain)
perm_test.l = permy(target_brain, ts_Ssec_liver, ts_protein_liver)

conf.int = function(x)
{
  se = sd(x)/sqrt(length(x))
  upper = mean(x) + (se * 1.96)
  lower = mean(x) - (se * 1.96)

  return(c(lower, upper))
}

#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
#write.csv(perm_test, "Permutation_Seldin_Report.csv")

@@

#setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
#perm_test = read.csv("Permutation_Seldin_Report.csv")
#perm_test= fix.read.csv(perm_test)

test_top_b = perm_test.b[rownames(perm_test.b) == top_brain[1],]
true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == top_brain[1],]

hist(as.numeric(test_top_b), main = "Null Ssec Distribution for Top Gene: Brain", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_top_b))[1], 2), round(conf.int(as.numeric(test_top_b))[2],2)), xlab = "Ssec Score", xlim = c(0, true$Ssec))
abline(v = true, col = "red")

test_bot_b = perm_test.b[rownames(perm_test.b) == bot_brain[1],]
true = ts_Ssec_brain[ts_Ssec_brain$Gene_symbol == bot_brain[1],]
hist(as.numeric(test_bot_b), main = "Null Ssec Distribution for Bottom Gene: Brain", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_bot_b))[1], 2), round(conf.int(as.numeric(test_bot_b))[2], 2)), xlab = "Ssec Score")
abline(v = true, col = "red")

test_top_l = perm_test.l[rownames(perm_test.l) == top_liver[1],]
true = ts_Ssec_liver[ts_Ssec_liver$Gene_symbol == top_liver[1],]
hist(as.numeric(test_top_l), main = "Null Ssec Distribution for Top Gene: Liver", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_top_l))[1], 2), round(conf.int(as.numeric(test_top_l))[2], 2)), xlab = "Ssec Score")
abline(v = true, col = "red")

test_bot_l = perm_test.l[rownames(perm_test.l) == bot_liver[1],]
true = ts_Ssec_liver[ts_Ssec_liver$Gene_symbol == bot_liver[1],]
hist(as.numeric(test_bot_l), main = "Null Ssec Distribution for Bottom Gene: Liver", sub = paste("Confidence Interval:", round(conf.int(as.numeric(test_bot_l))[1], 2), round(conf.int(as.numeric(test_bot_l))[2]), 2), xlab = "Ssec Score")
abline(v = true, col = "red")

## Selecting Relavent TES Protein-Coding Genes

Because our QQ plots are less than satisfactory, we will determine our top Ssec scores by selecting all TES protein-coding genes that have a higher Ssec score than the 95th percentile of Ssec scores for TENS protein-coding genes in the respective tissues. This results in the following six genes for brain and one gene for liver.

brain_quant = quantile(tsns_Ssec_brain$Ssec, probs = 0.95)
liver_quant = quantile(tsns_Ssec_liver$Ssec, probs = 0.95)

selected_brain = ts_Ssec_brain[ts_Ssec_brain$Ssec > brain_quant,]
selected_liver = ts_Ssec_liver[ts_Ssec_liver$Ssec > liver_quant,]

### Plots ###

# target genes

#setwd("C:/Users/angel/OneDrive/Desktop/Saba Lab/Data/")
setwd("C:/Users/yodeange/OneDrive - The University of Colorado Denver/Saba Lab/Data")
bicor.brain = read.csv("Bicor_brain.csv")
bicor.liver = read.csv("Bicor_liver.csv")
bicor.brain = fix.read.csv(bicor.brain)
bicor.liver = fix.read.csv(bicor.liver)

s.brain_1 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[1], ]
top.target.brain_1 = colnames(s.brain_1)[s.brain_1[1,] %in% sort(as.numeric(s.brain_1))[1:4]]

s.brain_2 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[2], ]
top.target.brain_2 = colnames(s.brain_2)[s.brain_2[1,] %in% sort(as.numeric(s.brain_2))[1:4]]

s.brain_3 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[3], ]
top.target.brain_3 = colnames(s.brain_3)[s.brain_3[1,] %in% sort(as.numeric(s.brain_3))[1:4]]

s.brain_4 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[4], ]
top.target.brain_4 = colnames(s.brain_4)[s.brain_4[1,] %in% sort(as.numeric(s.brain_4))[1:4]]

s.brain_5 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[5], ]
top.target.brain_5 = colnames(s.brain_5)[s.brain_5[1,] %in% sort(as.numeric(s.brain_5))[1:4]]

s.brain_6 = bicor.brain[rownames(bicor.brain) == selected_brain$Gene_symbol[6], ]
top.target.brain_6 = colnames(s.brain_6)[s.brain_6[1,] %in% sort(as.numeric(s.brain_6))[1:4]]

s.liver_1 = bicor.liver[rownames(bicor.liver) == selected_liver$Gene_symbol[1], ]
top.target.liver_1 = colnames(s.liver_1)[s.liver_1[1,] %in% sort(as.numeric(s.liver_1))[1:4]]

```{r, echo = FALSE}
## SUMMARY TABLE

names_brain = c("Cntn4", "Tnn", "Bmp3", "Wnt3", "Lamc3", "Cbln1")
names_liver = c("Kiss1")

s.brain_1top = s.brain_1[s.brain_1 < 0.05]
s.brain_2top = s.brain_2[s.brain_2 < 0.05]
s.brain_3top = s.brain_3[s.brain_3 < 0.05]
s.brain_4top = s.brain_4[s.brain_4 < 0.05]
s.brain_5top = s.brain_5[s.brain_5 < 0.05]
s.brain_6top = s.brain_6[s.brain_6 < 0.05]

numtarget_brain = c(s.brain_1top, s.brain_2top, s.brain_3top, s.brain_4top, s.brain_5top, s.brain_6top)

s.liver_1top = s.liver_1[s.liver_1 < 0.05]

fin_brain = data.frame(names_brain, selected_brain[,2], numtarget_brain)

fin_liver = data.frame(names_liver, selected_liver[,2], s.liver_1top)

knitr::kable(fin_brain, caption = "TES Protein-Coding Genes in the Brain that have higher Ssec scores than the 95th quantile of TENS protein-Coding Genes from the Brain")

knitr::kable(fin_liver, caption = "TES Protein-Coding Genes in the Liver that have higher Ssec scores than the 95th quantile of TENS Protein-Coding Genes fromt the Liver")

# plotting RNA expression

better.plot = function (protein, selected, target, top, tissue, not.tissue)
{
  par(mfrow = c(2,2))
  for (i in 1:4)
  {
    plot(protein[, colnames(protein) == selected], target[, colnames(target) == top[i]], xlab = paste("TES", tissue, selected), ylab = paste("Target", not.tissue, top[i]))
  }
}
```

To give a rough estimate of the influence these TES protein-coding genes have in the target tissue, we have plotted the expression of the top TES protein-coding gene against the expression of the four most significantly correlated target genes for each of our top genes. We hope to see defined slope that will indicate the influence these genes have.

### Highest Ssec Score TES Brain and its 4 most correlated liver genes

```{r, echo = FALSE}
better.plot(ts_protein_brain, selected_brain$Gene_symbol[1], target_liver, top.target.brain_1, "brain", "liver")

### Second Highest Ssec Score TES Brain and its 4 most correlated liver genes

```{r, echo = FALSE}
better.plot(ts_protein_brain, selected_brain$Gene_symbol[2], target_liver, top.target.brain_2, "brain", "liver")

### Third Highest Ssec Score TES Brain and its 4 most correlated liver genes

```{r, echo = FALSE}
better.plot(ts_protein_brain, selected_brain$Gene_symbol[3], target_liver, top.target.brain_3, "brain", "liver")

### Fourth Highest Ssec Score TES Brain and its 4 most correlated liver genes

```{r, echo = FALSE}
better.plot(ts_protein_brain, selected_brain$Gene_symbol[4], target_liver, top.target.brain_4, "brain", "liver")

### Fifth Highest Ssec Score TES Brain and its 4 most correlated liver genes

```{r, echo = FALSE}
better.plot(ts_protein_brain, selected_brain$Gene_symbol[5], target_liver, top.target.brain_5, "brain", "liver")

### Sixth Highest Ssec Score TES Brain and its 4 most correlated liver genes

```{r, echo = FALSE}
better.plot(ts_protein_brain, selected_brain$Gene_symbol[6], target_liver, top.target.brain_6, "brain", "liver")

### Highest Ssec Score TES Liver and its 4 most correlated brain genes

```{r, echo = FALSE}
better.plot(ts_protein_liver, selected_liver$Gene_symbol[1], target_brain, top.target.liver_1, "liver", "brain")

#### Appendix: TES v. All target tissue genes

# Calculate Ssec values: tissue-specific -> full
#### TS ####
## BRAIN ##
protein_liver = protein_liver[rownames(ts_protein_brain),]

protein_brain = protein_brain[rownames(ts_protein_liver),]

# Ssec distribtions TSNS v. TS, brain and liver
#Brain
h1 = cbind(rep("TENS brain", times = length(tsns_Ssec_brain$Ssec)), tsns_Ssec_brain$Ssec)
h2 = cbind(rep("TES brain", times = length(ts_Ssec_brain$Ssec)), ts_Ssec_brain$Ssec)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Brain Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")

# Liver
h1 = cbind(rep("TENS liver", times = length(tsns_Ssec_liver$Ssec)), tsns_Ssec_liver$Ssec)
h2 = cbind(rep("TES liver", times = length(ts_Ssec_liver$Ssec)), ts_Ssec_liver$Ssec)

ggplot(compare, aes(val, color = trt)) + 
  geom_density() +
  labs(title="Liver Ssec Distribution: Tissue-Exclusive Non-Secreted v. Tissue-Exclusive Secreted", x = "Ssec", y = "Density")

@@

@@

with(this, qqplot(val[trt == "Not Secreted"], val[trt == "Secreted"], xlab = "Not Secreted", ylab = "Secreted", main = "QQ Plot for Tissue-Exclusive Brain Proteins"))
abline(a = 0, b = 1)
